Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-01-06 paper.qmd
Updated Figure9

distance_points <- ggplot(
  distance_two_curvy_dist, 
  aes(x = distance,
      y = dist_highd)) +
  geom_point(alpha = 0.5) +
  #geom_jitter() +
  # scale_x_continuous(breaks = sort(unique(as.numeric(distance_two_curvy_dist$bin))),
  #                    labels = c(
  #                      expression(a[1]),
  #                      expression(2*a[1]),
  #                      expression(3*a[1]),
  #                      expression(4*a[1]),
  #                      expression(5*a[1]),
  #                      expression(6*a[1]),
  #                      expression(7*a[1]),
  #                      expression(8*a[1])
  #                    )) +
  # geom_vline(xintercept = 2,
  #            linetype="solid",
  #            color = "#bdbdbd",
  #            linewidth=1) +
  # geom_vline(xintercept = 4.5,
  #            linetype="solid",
  #            color = "#bdbdbd",
  #            linewidth=1) +
  ylab(expression(d^{(7)})) +
  xlab(expression(d^{(2)})) +
  #ggtitle("(a)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        panel.grid.major.x = element_blank(),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

@@

two_curvy_proj_umap_model <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") + #31a354
  geom_point(
    size = 0.5,
    alpha = 0.5,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.1, 0.1)) + 
  ylim(c(-0.1, 0.1)) +
  interior_annotation("c2", sc_ltr_pos) +
  theme(
    legend.position = "none"
  )

two_curvy_proj_umap_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") + #31a354
  geom_point(
    size = 0.5,
    alpha = 0.5,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.126, 0.125)) + 
  ylim(c(-0.126, 0.125)) +
  interior_annotation("a3", sc_ltr_pos) +
  theme(
    legend.position = "none"
  )

two_curvy_proj_umap_model_delaunay <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df1, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") + #31a354
  geom_point(
    size = 0.5,
    alpha = 0.5,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.126, 0.125)) + 
  ylim(c(-0.126, 0.125)) +
  interior_annotation("a2", sc_ltr_pos) +
  theme(
    legend.position = "none"
  )

two_curvy_proj_umap_model_benchmark2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df2, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") + #31a354
  geom_point(
    size = 0.5,
    alpha = 0.5,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.126, 0.125)) + 
  ylim(c(-0.126, 0.125)) +
  interior_annotation("b2", sc_ltr_pos) +
  theme(
    legend.position = "none"
  )

two_curvy_proj_umap_all_model <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = true_model_df_proj, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#2b8cbe") +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") +
  geom_point(
    size = 0.5,
    alpha = 0.5,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.126, 0.125)) + 
  ylim(c(-0.126, 0.125)) +
  interior_annotation("a4", sc_ltr_pos) +
  theme(
    legend.position = "none"
  )

>>>>

distance_points <- ggplot(
  distance_two_curvy_dist, 
  aes(x = distance,
      y = dist_highd)) +
  geom_point(alpha = 0.5) +
  #geom_jitter() +
  # scale_x_continuous(breaks = sort(unique(as.numeric(distance_two_curvy_dist$bin))),
  #                    labels = c(
  #                      expression(a[1]),
  #                      expression(2*a[1]),
  #                      expression(3*a[1]),
  #                      expression(4*a[1]),
  #                      expression(5*a[1]),
  #                      expression(6*a[1]),
  #                      expression(7*a[1]),
  #                      expression(8*a[1])
  #                    )) +
  geom_vline(xintercept = 2.5 * bin_width,
             linetype="solid",
             color = "#bdbdbd",
             linewidth=1) +
  geom_vline(xintercept = 2 * bin_width,
           linetype="solid",
           color = "#bdbdbd",
           linewidth=1) +
  # geom_vline(xintercept = 4.5,
  #            linetype="solid",
  #            color = "#bdbdbd",
  #            linewidth=1) +
  ylab(expression(d^{(7)})) +
  xlab(expression(d^{(2)})) +
  #ggtitle("(a)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        panel.grid.major.x = element_blank(),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

@@

two_curvy_proj_umap_model <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") + #31a354
  geom_point(
    size = 0.5,
    alpha = 0.2,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.1, 0.1)) + 
  ylim(c(-0.1, 0.1)) +
  interior_annotation("c2", sc_ltr_pos) +
  theme(
    legend.position = "none"
  )

two_curvy_proj_umap_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") + #31a354
  geom_point(
    size = 0.5,
    alpha = 0.2,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.1, 0.1)) + 
  ylim(c(-0.1, 0.1)) +
  interior_annotation("a3", sc_ltr_pos) +
  theme(
    legend.position = "none"
  )

two_curvy_proj_umap_model_delaunay <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df1, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") + #31a354
  geom_point(
    size = 0.5,
    alpha = 0.2,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.1, 0.1)) + 
  ylim(c(-0.1, 0.1)) +
  interior_annotation("a2", sc_ltr_pos) +
  theme(
    legend.position = "none"
  )

two_curvy_proj_umap_model_benchmark2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df2, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") + #31a354
  geom_point(
    size = 0.5,
    alpha = 0.2,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.1, 0.1)) + 
  ylim(c(-0.1, 0.1)) +
  interior_annotation("b2", sc_ltr_pos) +
  theme(
    legend.position = "none"
  )

two_curvy_proj_umap_all_model <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = true_model_df_proj, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#2b8cbe") +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") +
  geom_point(
    size = 0.5,
    alpha = 0.5,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.1, 0.1)) + 
  ylim(c(-0.1, 0.1)) +
  interior_annotation("a4", sc_ltr_pos) +
  theme(
    legend.position = "none"
  )
